---
title: "PredictingStocks_X"
author: "Hair Parra"
date: "May 10, 2020"
output:
  html_document:
    keep_md: yes
    keep_pdf: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.path = "../img/")
```

#  Historical Stocks Data Anlaysis: Forecasting Closing Prices

## Loading packages

```{r Library loading,warning=FALSE,message=FALSE}
library(tidyverse)
library(tidyquant)
library(gridExtra)
library(tibbletime)
library(forecast)
library(itsmr)
library(here)
library(bbmle)
library(tseries)
library(fpp2)
library(ggthemes)
library(readr)
library(xts)
library(reshape)
require(timeDate)
library(png)
knitr::opts_chunk$set(comment=NA,tidy=TRUE)
```

```{r, echo=FALSE}
# Create a new theme
theme_stonks <- function (base_size = 11, base_family = "") {
    theme_bw() %+replace% 
    theme(
      panel.grid.major  = element_line(), 
      panel.background = element_rect(fill = "lightblue"), # element_rect(fill = "black") 
      panel.border = element_rect(color = "lightblue", fill = NA),
      axis.line = element_line(color = "lightblue"),
      axis.ticks = element_line(color = "lightblue"),
      axis.text = element_text(color = "steelblue")
      )
}
```

## Loading the data 

```{r}
stocks_3M <- read_csv("../data_raw/stocks_data_3M.csv")
head(stocks_3M, 10)
```

## Data Preprocessing

Next, extract the columns of interest and convert into time series objects

```{r}
stocks_3M_data <- select(stocks_3M, Date, Close) # extract cols 
dates <- as.POSIXct.Date(stocks_3M_data$Date) # extract dates in POSIXct format
stocks_3M_data.ts <- xts(stocks_3M_data$Close, 
                        order.by = dates) # 7600
str(stocks_3M_data.ts) # inspect the data
```

## Inspecting the data


## Autoplot, ACF and PACF

```{r}
# Plot the same white noice this time as lines  
autoplot(stocks_3M_data.ts) + 
  geom_line(colour="blue")  +
  ggtitle("Stocks closing price historical data (3M)") + 
  theme_stonks() + xlab("Date") + ylab("USD") + geom_point(color="black")
```


```{r}
# ACF
ggAcf(stocks_3M_data.ts) + theme_stonks()
```

```{r}
# PACF 
ggPacf(stocks_3M_data.ts) + theme_stonks()
```


## Estimating the trend 

```{r}
# Estimate various trends
stocks_3M_linear <- tslm(ts(stocks_3M_data.ts)~trend)  
stocks_3M_p5 <- tslm(ts(stocks_3M_data.ts)~trend + I(trend^2) + I(trend^3) + I(trend^4) + I(trend^5) ) # polynomial
stocks_3M_ma5 <- ma(ts(stocks_3M_data.ts), order=5) # moving average
stocks_3M_trends <- data.frame(cbind(Data=stocks_3M_data.ts,  # stack in a dataframe
                        Linear_trend=fitted(stocks_3M_linear),
                        Poly_trend=fitted(stocks_3M_p5),
                        Moving_avg5 = stocks_3M_ma5
                        ))

# transform to xts objects
stocks_3M_linear <- xts(fitted(stocks_3M_linear), order.by = dates)
stocks_3M_p5 <- xts(fitted(stocks_3M_p5), order.by = dates)

# Plot all the trends together 
autoplot(stocks_3M_data.ts, colour="original") + theme_stonks() + 
  geom_line(aes(y=stocks_3M_linear, color="linear"),size=1) + 
  geom_line(aes(y=stocks_3M_p5, color = "O(5) poly"), size=1) + 
  geom_line(aes(y=stocks_3M_ma5, color ="ma21"), size=1)  + 
  scale_color_manual(values = c('original'= 'blue', 
                                'linear' = 'darkblue',
                                'O(5) poly' = 'red', 
                                'ma21'= 'yellow')) + 
  labs(color = 'Trend fit') +  ylab("USD") + 
  ggtitle("Different trend fits for the stocks data") 
```

```{r}
# Detrend and show the de-trended series
stocks_3M_ma21 <- xts(stocks_3M_p5,order.by = dates) # cast to xts 
detrend_stocks_3M <- stocks_3M_data.ts - stocks_3M_p5 # substract from original

# Plot the residuals
autoplot(detrend_stocks_3M) + theme_stonks() + 
  ggtitle("De-trended Data ( O(5) trend)") +
  geom_hline(yintercept = 0, colour="black") + 
  geom_point() + ylab("USD - trend")  + geom_line(color="blue")
```

The residuals look zero-trended. 

```{r}
# ACF
ggAcf(detrend_stocks_3M) + theme_stonks() 
```

The ACF lags all , except for one fall within the 0.25 confidence bounds. 

```{r}
# PACF 
ggPacf(detrend_stocks_3M) + theme_stonks() 
```

The PACF residuals mostly fall within the confidence bounds; whoever there seems to be some negative autocorrelation present across lags. However, from all the previous, there doesn't seem to be a strong seasonal component present.

## Train-test split & ARIMA fitting

We will now split the data into 32 training data points and 10 test data points. We will produce predictions and compare them to assess fit.

```{r}
## train_test_split
detrend_stocks_3M_train <- stocks_3M_data.ts[1:(round(length(detrend_stocks_3M))-10)] # 32
detrend_stocks_3M_test <- stocks_3M_data.ts[(round(length(detrend_stocks_3M))-9):length(detrend_stocks_3M)] # 10
str(detrend_stocks_3M_train)
str(detrend_stocks_3M_test)
length(detrend_stocks_3M_train)
length(detrend_stocks_3M_test)
```

```{r}
# Fit the ARIMA model on trian data
detrend_stocks_3M_arima_110 = auto.arima(detrend_stocks_3M_train,
                       seasonal=TRUE,
                       stepwise=FALSE,
                       max.d = 2, 
                       ic = c("aicc", "aic", "bic") , 
                       approximation=FALSE,
                       trace=TRUE)
```

```{r}
detrend_stocks_3M_arima_110
```

## Inspecting the residuals

```{r}
checkresiduals(detrend_stocks_3M_arima_110) 
```

```{r}
# check the residuals 
train_dates <- as.POSIXct.Date(stocks_3M_data$Date[1: (length(stocks_3M_data$Date) - 10 )])
arima_res <- xts(residuals(detrend_stocks_3M_arima_110), 
                 order.by = train_dates)
```


```{r}
# Plot the residuals
autoplot(arima_res) + theme_stonks() + 
  geom_point() + geom_line(color="blue") + 
  geom_hline(yintercept = 0, colour="black") 
```

We see that perhaps around Mars 13, there could have been a possible outlier. 

```{r}
# ACF 
ggAcf(residuals(detrend_stocks_3M_arima_110)) + theme_stonks() 
```

```{r}
# PACF 
ggPacf(residuals(detrend_stocks_3M_arima_110)) + theme_stonks() 
```

In both cases, the ACF and PACF points find whithin confidence bounds, with the exception of one. This one might be due to the possible utlier we had before. 


```{r}
# Inspect roots
autoplot(detrend_stocks_3M_arima_110) + theme_stonks() 
```

The roots of the AR(1) polynomial guarantee the process is stationary and causal, and of course, it is also invertible. We can also verify this by performing the ADF and KPSS tests for stationarity: 

```{r}
# Test with a bunch of different k's ? (bigger augmented versions)
detrend_stocks_3M_arima_110_diff <- diff(residuals(detrend_stocks_3M_arima_110), lag=1) # difference order 1
adf.test(detrend_stocks_3M_arima_110_diff,k=1) # ADF
kpss.test(detrend_stocks_3M_arima_110_diff) # KPSS
```

Reject -> stationary for the ADF, fail to reject -> stationary for the KPSS. Now we can proceed with the forecasting. 


## Forecasting


### Obtaining model and trend forecasts 

We will now forecast 10 observations from both the main model and the trend

```{r}
detrend_stocks_3M_arima_110_forecasts <- forecast::forecast(detrend_stocks_3M_arima_110,h=10) # ARIMA(1,1,0) forecasts
forecasted_trend <- forecast::forecast( stocks_3M_p5, h=10)  # forecast 10 trend observations
```


### Model forecasts

Let's produce a table with the point forecast values along with the errors and confidence intervals for predictions

```{r, echo=FALSE}
# Print the forecasts along with the boundaries 
forecast_table<-print(detrend_stocks_3M_arima_110_forecasts) %>%
  mutate(observed=detrend_stocks_3M_test,
         errors=`Point Forecast`-observed)
```

```{r}
# Show the table with errors 
forecast_table = as.data.frame(forecast_table)
colnames(forecast_table) <- c("Point_Forecast","Lo80","Hi80","Lo95","Hi95","observed","errors")
forecast_table
```

Extract the values as plain vectors for plotting: we paste this to a bunch of `NA` values to be able to plot all together. 

```{r}
predicts <- c(rep(NA,32),forecast_table$Point_Forecast)
predicts_Lo80 <-  c(rep(NA,32),forecast_table$Lo80)
predicts_Hi80 <-  c(rep(NA,32),forecast_table$Hi80)
predicts_Lo95 <-  c(rep(NA,32),forecast_table$Lo95)
predicts_Hi95 <-  c(rep(NA,32),forecast_table$Hi95)
```


### Producing the forecasts 

```{r}
# Plot the predictions + xlim(1.05e+08,1.09e+08) + ylim(32,45) 
autoplot(detrend_stocks_3M_arima_110_forecasts)  + theme_stonks()
```

Although this plot looks pretty, notice the scale is somehow very off! We will construct a better plot manually, by using the values we obtained before, so that we can see both the original data and the repdictions along with the confidence intervals like above. 

```{r}
autoplot(stocks_3M_data.ts, colour="orig") + theme_stonks() + 
  geom_line(aes(y=predicts,colour = "predicted") ) + 
   geom_ribbon(aes(x=as.POSIXct.Date(stocks_3M_data$Date),
                   ymin=predicts_Lo95,ymax=predicts_Hi95),fill="pink", alpha=.3) + 
   geom_ribbon(aes(x=as.POSIXct.Date(stocks_3M_data$Date),
                   ymin=predicts_Lo80,ymax=predicts_Hi80),fill="blue", alpha=.5) + 
    scale_color_manual(values = c('predicted'= 'red',
                                  'orig'='black')
                       ) + 
  ylab("USD") + xlab("Date") + geom_point() + geom_line(color="blue") + 
  labs(color = 'Trend fit') + ggtitle("Stocks predictions for 10 days")
```




















